19Q1

### 2019-05-06 19Q1.Rmd

WilcoxonTestDF <- function(dataframe, grouping, outfile) {
  require(ggpubr)
  if(grouping == "Drug_Gene_Lin") {
    # Filter out DGLs with too few samples in at least one mutation status category
    count <- ctrp_ptmuts %>% count(Drug_Gene_Lin, Mutation_Status)
    count2 <- count %>% count(Drug_Gene_Lin)
    count3 <- merge(count, count2, by = "Drug_Gene_Lin", suffixes = c("_sub3", "_1mut")) %>% filter(n_1mut == 2)
    count3$n_sub3 <- ifelse(count3$n_sub3 >= 3, TRUE, FALSE)
    count4 <- count3 %>% filter(n_sub3) %>% count(Drug_Gene_Lin) %>% filter(n == 2)
    dataframe <- filter(dataframe, Drug_Gene_Lin %in% unique(count4$Drug_Gene_Lin))
    dataframe$Drug_Gene_Lin <- factor(dataframe$Drug_Gene_Lin)
  }
  signif <- compare_means(Score ~ Mutation_Status, group.by = grouping, data = dataframe, paired = FALSE, method = "wilcox.test", p.adjust.method = "BH")
  signif <- adj_signif(signif, alpha = 0.05)
  signif <- signif[order(signif$p),]
  if(grouping == "Drug_Gene") {
    signif$InG2P <- ifelse(signif$Drug_Gene %in% g2p_druggenes, "Yes", "No")
  }
  if(grouping == "Drug_Gene_Lin") {
    signif$InG2P <- ifelse(signif$Drug_Gene_Lin %in% g2p_dgls, "Yes", "No")
  }
  saveRDS(signif, outfile)
  return(signif)
}
```

```{r wilcoxon_tests, eval=FALSE}
crispr_signif_gene <- WilcoxonTestDF(dataframe = crispr_ptmuts, grouping = "Gene", outfile = "./data_munging/rds/crispr_signif_g2p_gene.rds")
crispr_signif_lineage <- WilcoxonTestDF(dataframe = crispr_ptmuts, grouping = c("Gene", "Primary.Disease"), outfile = "./data_munging/rds/crispr_signif_g2p_lineage.rds")

ccle_signif_gene <- WilcoxonTestDF(dataframe = ccle_ptmuts, grouping = "Drug_Gene", outfile = "./data_munging/rds/ccle_signif_g2p_gene.rds")
ccle_signif_g2p_druggene <- WilcoxonTestDF(dataframe = filter(ccle_ptmuts, Drug %in% g2p_drugs & Gene %in% g2p_genes), grouping = "Drug_Gene", outfile = "./data_munging/rds/ccle_signif_g2p_druggene.rds")
ccle_signif_lineage <- WilcoxonTestDF(dataframe = ccle_ptmuts, grouping = "Drug_Gene_Lin", outfile = "./data_munging/rds/ccle_signif_g2p_lineage.rds")

ctrp_signif_gene <- WilcoxonTestDF(dataframe = ctrp_ptmuts, grouping = "Drug_Gene", outfile = "./data_munging/rds/ctrp_signif_g2p_gene.rds")
ctrp_signif_g2p_druggene <- WilcoxonTestDF(dataframe = filter(ctrp_ptmuts, Drug %in% g2p_drugs & Gene %in% g2p_genes), grouping = "Drug_Gene", outfile = "./data_munging/rds/ctrp_signif_g2p_druggene.rds")
ctrp_signif_lineage <- WilcoxonTestDF(dataframe = ctrp_ptmuts, grouping = "Drug_Gene_Lin", outfile = "./data_munging/rds/ctrp_signif_g2p_lineage.rds")

gdsc_signif_gene <- WilcoxonTestDF(dataframe = gdsc_ptmuts, grouping = "Drug_Gene", outfile = "./data_munging/rds/gdsc_signif_g2p_gene.rds")
gdsc_signif_g2p_druggene <- WilcoxonTestDF(dataframe = filter(gdsc_ptmuts, Drug %in% g2p_drugs & Gene %in% g2p_genes), grouping = "Drug_Gene", outfile = "./data_munging/rds/gdsc_signif_g2p_druggene.rds")
gdsc_signif_lineage <- WilcoxonTestDF(dataframe = gdsc_ptmuts, grouping = "Drug_Gene_Lin", outfile = "./data_munging/rds/gdsc_signif_g2p_lineage.rds")
```

@@

crispr_gene_mut_plot <- ggplot(data = crispr_ptmuts) +
  facet_wrap(~ Gene, nrow = 2, drop = FALSE) +
  scale_color_manual(name = "Mutation Status", labels = c("Wildtype", "Mutant"), values = c("1" = "darkorchid", "0" = "thistle")) +
  geom_jitter(aes(x = Mutation_Status, y = Score, color = Mutation_Status), width = 0.2, alpha = 0.5, size = 0.6) +
  geom_boxplot(aes(x = Mutation_Status, y = Score), fill = "transparent", outlier.shape = NA, lwd = 0.3) +
  geom_text(data = crispr_signif_gene, aes(x = 1.5, y = 1.75, label = p.signif.adj)) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  theme(legend.position = "top", legend.direction = "horizontal") +
  guides(color = guide_legend(override.aes = list(size = 3))) +
  labs(y = "CERES Score")
crispr_gene_mut_plot

@@

```{r heatmap_function}
MakeMutHeatmap <- function(signif_g2p_druggene, row_names, col_title, color, legend_title) {
  signif_g2p_druggene$Drug <- sapply(strsplit(signif_g2p_druggene$Drug_Gene, "_", fixed = TRUE), function(x) x[1])
  signif_g2p_druggene$Gene <- sapply(strsplit(signif_g2p_druggene$Drug_Gene, "_", fixed = TRUE), function(x) x[2])

@@

## Drugscreens

@@

# CCLE
ccle_cor_res <- ccle_ptmuts %>% group_by(Drug_Gene) %>%summarize(
                GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate,
                GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value,
                CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate,
                CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value)
ccle_cor_res <- merge(g2p_druggene_tally, ccle_cor_res, by = "Drug_Gene", all.y = TRUE)
ccle_cor_res$n <- ifelse(is.na(ccle_cor_res$n), 0, ccle_cor_res$n)
ccle_cor_res$InG2P <- ifelse(ccle_cor_res$n != 0, "Yes", "No")
ccle_cor_res$Dataset <- "CCLE"

# CTRP
ctrp_cor_res <- ctrp_ptmuts %>% group_by(Drug_Gene) %>%summarize(
                GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate,
                GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value,
                CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate,
                CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value)
ctrp_cor_res <- merge(g2p_druggene_tally, ctrp_cor_res, by = "Drug_Gene", all.y = TRUE)
ctrp_cor_res$n <- ifelse(is.na(ctrp_cor_res$n), 0, ctrp_cor_res$n)
ctrp_cor_res$InG2P <- ifelse(ctrp_cor_res$n != 0, "Yes", "No")
ctrp_cor_res$Dataset <- "CTRP"

# GDSC
gdsc_cor_res <- gdsc_ptmuts %>% group_by(Drug_Gene) %>%summarize(
                GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate,
                GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value,
                CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate,
                CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value)
gdsc_cor_res <- merge(g2p_druggene_tally, gdsc_cor_res, by = "Drug_Gene", all.y = TRUE)
gdsc_cor_res$n <- ifelse(is.na(gdsc_cor_res$n), 0, gdsc_cor_res$n)
gdsc_cor_res$InG2P <- ifelse(gdsc_cor_res$n != 0, "Yes", "No")
gdsc_cor_res$Dataset <- "GDSC"

# All
g2p_cor <- rbind(ccle_cor_res, ctrp_cor_res, gdsc_cor_res)
```

@@

@@

Kolmogorov-Smirnov tests comparing DGLs that are in a G2P DGL and not.
```{r drug_KS_tests, eval=TRUE, warning=FALSE, fig.height=6, fig.width=12}
# g2p_corr_all <- G2PCorrTable(ccle_ptmuts_df = ccle_ptmuts, ctrp_ptmuts_df = ctrp_ptmuts, gdsc_ptmuts_df = gdsc_ptmuts)
# saveRDS(g2p_corr_all, "./data_munging/rds/g2p_corr_all.rds")
g2p_corr_all <- readRDS("./data_munging/rds/g2p_corr_all.rds")
g2p_mut_stats <- G2PCorrStatsTables(corr_df = g2p_corr_all, table_type = "mutation")

>>>>

WilcoxonTestDF <- function(dataframe, grouping, outfile) {
  require(ggpubr)
  if("Drug_Gene_Lin" %in% grouping) {
    # Filter out DGLs with too few samples in at least one mutation status category
    count_mut_n <- dataframe %>% count(Drug_Gene_Lin, Mutation_Status) %>% filter(n >= 3)
    count_mut_category <- count_mut_n %>% count(Drug_Gene_Lin) %>% filter(n == 2)
    dgl_keep <- intersect(count_mut_n$Drug_Gene_Lin, count_mut_category$Drug_Gene_Lin)
    dataframe <- filter(dataframe, Drug_Gene_Lin %in% dgl_keep)
    dataframe$Drug_Gene_Lin <- as.character(dataframe$Drug_Gene_Lin)
  }
  signif <- compare_means(Score ~ Mutation_Status, group.by = grouping, data = dataframe, paired = FALSE, method = "wilcox.test", p.adjust.method = "BH")
  signif <- adj_signif(signif, alpha = 0.05)
  signif <- signif[order(signif$p),]
  if("Drug_Gene" %in% grouping) {
    signif$InG2P <- ifelse(signif$Drug_Gene %in% g2p_druggenes, "Yes", "No")
  }
  if("Drug_Gene_Lin" %in% grouping) {
    signif$InG2P <- ifelse(signif$Drug_Gene_Lin %in% g2p_dgls, "Yes", "No")
  }
  saveRDS(signif, outfile)
  return(signif)
}
```

```{r wilcoxon_tests, eval=FALSE}
crispr_ptmuts <- unfactorize(crispr_ptmuts)
crispr_signif_gene <- WilcoxonTestDF(dataframe = crispr_ptmuts, grouping = "Gene", outfile = "./data_munging/rds/crispr_signif_g2p_gene.rds")
crispr_signif_lineage <- WilcoxonTestDF(dataframe = crispr_ptmuts, grouping = c("Gene", "Primary.Disease"), outfile = "./data_munging/rds/crispr_signif_g2p_lineage.rds")

ccle_ptmuts_filt <- filter(ccle_ptmuts, Drug %in% g2p_drugs)
ccle_ptmuts_filt$Gene <- factor(ccle_ptmuts_filt$Gene, ordered = FALSE)
ccle_ptmuts_filt <- unfactorize(ccle_ptmuts_filt)
ccle_signif_gene <- WilcoxonTestDF(dataframe = ccle_ptmuts, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/ccle_signif_g2p_gene.rds")
ccle_signif_g2p_druggene <- WilcoxonTestDF(dataframe = filter(ccle_ptmuts, Drug %in% g2p_drugs & Gene %in% g2p_genes), grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/ccle_signif_g2p_druggene.rds")
ccle_signif_lineage <- WilcoxonTestDF(dataframe = ccle_ptmuts_filt, grouping = c("Drug", "Gene", "Primary.Disease", "Drug_Gene_Lin"), outfile = "./data_munging/rds/ccle_signif_g2p_lineage.rds")

ctrp_ptmuts_filt <- filter(ctrp_ptmuts, Drug %in% g2p_drugs)
ctrp_ptmuts_filt$Gene <- factor(ctrp_ptmuts_filt$Gene, ordered = FALSE)
ctrp_ptmuts_filt <- unfactorize(ctrp_ptmuts_filt)
ctrp_signif_gene <- WilcoxonTestDF(dataframe = ctrp_ptmuts, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/ctrp_signif_g2p_gene.rds")
ctrp_signif_g2p_druggene <- WilcoxonTestDF(dataframe = filter(ctrp_ptmuts, Drug %in% g2p_drugs & Gene %in% g2p_genes), grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/ctrp_signif_g2p_druggene.rds")
ctrp_signif_lineage <- WilcoxonTestDF(dataframe = ctrp_ptmuts_filt, grouping = c("Drug", "Gene", "Primary.Disease", "Drug_Gene_Lin"), outfile = "./data_munging/rds/ctrp_signif_g2p_lineage.rds")

gdsc_ptmuts_filt <- filter(gdsc_ptmuts, Drug %in% g2p_drugs)
gdsc_ptmuts_filt$Gene <- factor(gdsc_ptmuts_filt$Gene, ordered = FALSE)
gdsc_ptmuts_filt <- unfactorize(gdsc_ptmuts_filt)
gdsc_signif_gene <- WilcoxonTestDF(dataframe = gdsc_ptmuts, grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/gdsc_signif_g2p_gene.rds")
gdsc_signsif_g2p_druggene <- WilcoxonTestDF(dataframe = filter(gdsc_ptmuts, Drug %in% g2p_drugs & Gene %in% g2p_genes), grouping = c("Drug", "Gene", "Drug_Gene"), outfile = "./data_munging/rds/gdsc_signif_g2p_druggene.rds")
gdsc_signif_lineage <- WilcoxonTestDF(dataframe = gdsc_ptmuts_filt, grouping = c("Drug", "Gene", "Primary.Disease", "Drug_Gene_Lin"), outfile = "./data_munging/rds/gdsc_signif_g2p_lineage.rds")
```

@@

crispr_gene_mut_plot <- ggplot(data = crispr_ptmuts) +
  facet_wrap(~ Gene, nrow = 2, drop = FALSE) +
  scale_color_manual(name = "Mutation Status", labels = c("Wildtype", "Mutant"), values = c("1" = "darkorchid", "0" = "thistle")) +
  scale_x_discrete(name = "Mutation Status", labels = c("0" = "WT", "1" = "Mut")) +
  geom_jitter(aes(x = Mutation_Status, y = Score, color = Mutation_Status), width = 0.2, alpha = 0.5, size = 0.6) +
  geom_boxplot(aes(x = Mutation_Status, y = Score), fill = "transparent", outlier.shape = NA, lwd = 0.3) +
  geom_text(data = crispr_signif_gene, aes(x = 1.5, y = 1.75, label = p.signif.adj)) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  theme(legend.position = "top", legend.direction = "horizontal") +
  guides(color = guide_legend(override.aes = list(size = 3))) +
  labs(y = "CERES Score")
# crispr_gene_mut_plot

@@

```{r mut_heatmap_function}
MakeMutHeatmap <- function(signif_g2p_druggene, row_names, col_title, color, legend_title) {
  # signif_g2p_druggene$Drug <- sapply(strsplit(signif_g2p_druggene$Drug_Gene, "_", fixed = TRUE), function(x) x[1])
  # signif_g2p_druggene$Gene <- sapply(strsplit(signif_g2p_druggene$Drug_Gene, "_", fixed = TRUE), function(x) x[2])

@@

@@

# CCLE
ccle_cor_res <- ccle_ptmuts %>% group_by(Drug, Gene, Drug_Gene) %>% summarize(
                GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate,
                GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value,
                CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate,
                CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value)
ccle_cor_res <- merge(g2p_druggene_tally, ccle_cor_res, by = "Drug_Gene", all.y = TRUE)
ccle_cor_res$n <- ifelse(is.na(ccle_cor_res$n), 0, ccle_cor_res$n)
ccle_cor_res$InG2P <- ifelse(ccle_cor_res$n != 0, "Yes", "No")
ccle_cor_res$Dataset <- "CCLE"

# CTRP
ctrp_cor_res <- ctrp_ptmuts %>% group_by(Drug, Gene, Drug_Gene) %>% summarize(
                GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate,
                GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value,
                CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate,
                CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value)
ctrp_cor_res <- merge(g2p_druggene_tally, ctrp_cor_res, by = "Drug_Gene", all.y = TRUE)
ctrp_cor_res$n <- ifelse(is.na(ctrp_cor_res$n), 0, ctrp_cor_res$n)
ctrp_cor_res$InG2P <- ifelse(ctrp_cor_res$n != 0, "Yes", "No")
ctrp_cor_res$Dataset <- "CTRP"

# GDSC
gdsc_cor_res <- gdsc_ptmuts %>% group_by(Drug, Gene, Drug_Gene) %>% summarize(
                GE_r = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$estimate,
                GE_p = cor.test(y = Score, x = TPM, method = "spearman", use = "complete.obs")$p.value,
                CN_r = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate,
                CN_p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value)
gdsc_cor_res <- merge(g2p_druggene_tally, gdsc_cor_res, by = "Drug_Gene", all.y = TRUE)
gdsc_cor_res$n <- ifelse(is.na(gdsc_cor_res$n), 0, gdsc_cor_res$n)
gdsc_cor_res$InG2P <- ifelse(gdsc_cor_res$n != 0, "Yes", "No")
gdsc_cor_res$Dataset <- "GDSC"

# All
g2p_cor <- rbind(ccle_cor_res, ctrp_cor_res, gdsc_cor_res)

saveRDS(g2p_cor, "./data_munging/rds/drug_cor.rds")
```

```{r load_drug_spear}
g2p_cor <- readRDS("./data_munging/rds/drug_cor.rds")
```

@@

```{r corr_heatmap_function}
MakeCorrHeatmap <- function(drug_cor_res, row_names, col_title, color, legend_title) {
  # drug_cor_res$Drug <- sapply(strsplit(signif_g2p_druggene$Drug_Gene, "_", fixed = TRUE), function(x) x[1])
  # drug_cor_res$Gene <- sapply(strsplit(signif_g2p_druggene$Drug_Gene, "_", fixed = TRUE), function(x) x[2])

  # signif_grid_annot <- dcast(signif_g2p_druggene, Gene ~ Drug, value.var = "InG2P", fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
  # signif_grid_annot <- signif_grid_annot[g2p_gene_order, intersect(g2p_drug_order, colnames(signif_grid))]
  # signif_grid_annot <- signif_grid_annot[complete.cases(signif_grid_annot), ]
  # signif_grid_annot[signif_grid_annot == "No"] <- ""
  # signif_grid_annot[signif_grid_annot == "Yes"] <- "*"

  col_fun <- colorRamp2(c(0, 0.05), c(color, "gray97"))
  # Legend properties list
  ht_legend_list <- list(title = legend_title, title_position = "topleft", legend_direction = "horizontal", legend_width = unit(2, "in"), border = "gray70",   at = seq(0, 0.05, by = 0.01), labels = seq(0, 0.05, by = 0.01))
  ht <- Heatmap(signif_grid,
                col = col_fun,
                na_col = "gray97",
                rect_gp = gpar(col = "gray70"),
                heatmap_legend_param = ht_legend_list,
                cluster_rows = FALSE, 
                cluster_columns = FALSE, 
                show_row_names = as.logical(row_names),
                row_names_side = "left",
                row_names_gp = gpar(fontsize = 10),
                column_names_side = "bottom",
                column_names_gp = gpar(fontsize = 10),
                show_column_names = FALSE,
                bottom_annotation = HeatmapAnnotation(
                  text = anno_text(colnames(signif_grid),
                                   rot = 45,
                                   gp = gpar(fontsize = 11),
                                   offset = unit(1, "npc"),
                                   just = "right"),
                  annotation_height = unit(0.85, "in")),
                  # annotation_height = max_text_width(colnames(signif_grid))),
                row_title = "Genes",
                row_title_gp = gpar(fontface = "bold", fontsize = 14),
                column_title = col_title, 
                column_title_gp = gpar(fontface = "bold", fontsize = 14),
                column_title_side = "bottom",
                # cell_fun = function(j, i, x, y, width, height, fill) { grid.text(signif_grid_annot[i, j], x, y, gp = gpar(fontsize = 20)) }
                )
  return(ht)
}
```

```{r drug_spear_plot}

```

@@

Kolmogorov-Smirnov tests comparing DGLs that are in a G2P DGL and not.
```{r drug_KS_tests, eval=TRUE, warning=FALSE, fig.height=6, fig.width=12}
# g2p_corr_all <- G2PCorrTable(ccle_ptmuts_df = ccle_ptmuts, ctrp_ptmuts_df = ctrp_ptmuts, gdsc_ptmuts_df = gdsc_ptmuts)
# saveRDS(g2p_corr_all, "./data_munging/rds/g2p_dgl_corr.rds")
g2p_corr_all <- readRDS("./data_munging/rds/g2p_dgl_corr.rds")
g2p_mut_stats <- G2PCorrStatsTables(corr_df = g2p_corr_all, table_type = "mutation")
